### 2016-02-08 text/MS.Rmd
implemented LOO model selection and updated text, tables, figures, numbers and functions

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
load('../output/RDatafiles/analyses.RData')
load('../output/RDatafiles/numbers.RData')
source('../R/functions-analyses.R')
source('../R/functions-figures.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=TRUE)
```
**Temperature effects on mass-scaling exponent of metabolic rates: a manipulative test**

*draft date: `r as.character(Sys.Date())`*

Diego R. Barneche, Craig R. White, Dustin J. Marshall

Centre for Geometric Biology / School of Biological Sciences, Monash University, Clayton, Victoria, Australia 3800

**Summary**

1. Organismal metabolic rate often scales allometrically with body mass, both inter and intra-specifically. Metabolic rates are also known to increase with temperature, but most theoretical tests of mass scaling and temperature dependence have treated both phenomena independently.  
2. In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species belonging to two phyla of colonial marine invertebrates.  
3. We show that mass-scaling exponents and temperature dependencies interact in different ways for different species. In so doing, we discuss the potentials of our results to multi-level scaling theory, showing that some species might be winners while others might be losers under different temperature scenarios.  
4. These findings highlight the need for new metabolic theory that formally incorporate interactions between mass-scaling and temperature dependence in ectothermic organisms, in order to better explain documented variation in the mass-scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.

**Key-words:** energetics, scaling, thermal limits, latitude, metabolic theory

**Introduction**

Metabolic theories provide a potential to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`. Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`. The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent. Numerous studies have demonstrated that, for most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates varying with many biotic and abiotic variables (e.g. `r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1, and averaging around 0.75. While these studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size. Such correlative studies cannot address the causal association between $B_i$ and $M_i$, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g. `r myCite(c('peters1983', 'calder1984book', 'schmidtnielsen1984'))` for comprehensive reviews) and many of these variables are also associated with $B_i$ `r myCite(c('mcnab2002book', 'konarzewski2013jcompphysb', 'white2013jcompphysb'))`.

In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for more definitive tests of metabolic theories. Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of body size and life-history traits (e.g. `r myCite('white2011amnat')`). Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`. For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (but see `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'price2007pnas', 'koontz2009amjbot'))`). Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes affect metabolic rates. For instance, it has been demonstrated that mimicking partial predation through size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`. However, to our knowledge, few studies have explicitly investigated the effects of partial predation on individual metabolic rates (e.g. `r myCite('white2011amnat')`).

Metabolic rates are also influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`. Existing metabolic theories assume nevertheless different paths through which temperature affects metabolic rates. For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` assumes that temperature affects the metabolic normalisation only `r myCite(c('gillooly2001', 'brown2004'))`, i.e. mass-scaling and temperature dependence of metabolic rates are fundamentally independent; contrarily, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`. Whether temperature dependence and mass scaling truly interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`. Testing such interaction is of great ecological significance because different mass-scaling exponents under different temperatures would potentially indicate contrasting life-history trajectories `r myCite('sebens1987annurevecolsys')` for a given species under different environments **(REFS??)**.

In the present study we experimentally manipulate size in different colonial marine organisms from two phyla (Bryozoa [Ectoprocta] and Porifera) in order to test whether the mass scaling and temperature dependence of metabolic rates interact `r myCite(c('white2011amnat', 'kearney2012amnat'))`. In doing so, we explicitly test if these interactions are general or species specific using a Bayesian model selection procedure. Finally, we discuss the relevance of our results to the life history of different species living under different environments.

**Material and Methods**

*1. Animal collections*  
To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B. stolonifera* (arborescent bryozoan), *Hippopodina* sp. (encrusting bryozoan), and an unidentified species of sponge belonging to the Microcionidae family `r myCite('hart2009ecology')` (Table 1). After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements. For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see `r myCite('white2011amnat')` for details and tests of potential experimental artefacts). Arborescent colonies were cut directly from the acetate sheets and fragmented.

*2. Respiration measurements*  
Individual metabolic rates (**provide units**, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`. $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:

$$\dot{V}\textrm{O}_\textrm{2} = ‑1 \left(\frac{m_a – m_b}{100}\right) V~{\beta}_{\textrm{O}_\textrm{2}}$$, (1)

where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) [**double-check these numbers. Shouldn't the second be 25 ˚C?**] `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals). At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.). Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany). All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.

*3. Statistical analyses*  
To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data

$$\textrm{ln}B_i \sim \textrm{ln}M_i * Temperature * Species$$, (2)

where mass (mg), $\textrm{ln}M_i$, is a continuous variable, while temperature (two levels: 10 ˚C and 25 ˚C) and species (four levels corresponding to the four species) are categorical variables. With this model, we allow each species to have a different mass-scaling slope that may change with temperature; we also account for species- and temperature-specific deviations on the metabolic normalisation (i.e. statistical model). We also included date of experiment, $\Delta D$ as a random effect on the metabolic normalisation to help control for natural experimental variability. We then performed a Bayesian model selection procedure (see below) by comparing the above full model with a series of nested two- and one-way interaction models in order to formally test for mass-temperature interactions in metabolic rates among different species.

Both models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters. Random effects were assumed to be normally distributed, with means of 0. Fitted parameters were assigned priors that were vague (i.e. locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing five chains of 1,000,000 steps each, including 500,000-step burn-in periods. Chains were thinned using a 250-step interval, so a total of 10,000 steps were retained to estimate posterior distributions (i.e. 5 $\times$ (1,000,000 - 500,000)/250 = 10,000). Each nested model was compared against the full model using leave-one-out cross-validation (LOO), which, similar to widely applicable information criterion (WAIC), is a fully Bayesian model selection procedure for estimating pointwise out-of-sample prediction accuracy using the log-likelihood evaluated at the posterior simulations of the parameter values `r myCite('hooten2015ecology')`. Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models, and models were considered significantly different at $\alpha \le$ 0.05. Model comparison was implemented using the R package *loo* version 0.1.3 `r myCite('loo')`. All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.

**Results**

Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1). Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e. intercepts), $B_o$, at higher temperatures (though not statistically significant for Microcionidae). For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B. stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`. It is worth noting that this $Q_{10}$ value is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV (i.e. $E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite(c('gillooly2001','yvondurocher2012a'))`. Different from the above species, the best model indicates that *Bugula neritina* and *Hippopodina* sp. present opposing interactions between mass-scaling and temperature dependence (Figure 1). For example, average posterior parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp. decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.

Posterior distributions for mass-scaling exponents show varied consistency with predictions from the two metabolic theories. For example, *Hippopodina* sp. shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`). On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`). Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`). Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).

**Discussion**

Here, we show experimentally that the mass-scaling of metabolic rates changes between two extreme temperatures, and in different ways across different species of marine invertebrates. We build on recent empirical evidence documenting similar changes (e.g. `r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates. Our results, however, do not indicate consistent mass-temperature interactions across growth forms. Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes (e.g. `r myCite('rezende2014functecol')`). Our findings are constrained to acute-related changes, and therefore do not consider temporal acclimation or even adaptation to changes in environmental temperature.

The fact that $B_o$ and $\alpha$ covary in different ways depending on species and temperature (Figure 1) suggests that the metabolic costs of maintenance will vary for different species at different environmental temperatures. For example, *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures, possibly indicating that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments. Thus, for this species, all else being equal, increases in temperature decrease the metabolic benefits of attaining a larger size while decreases in temperature reduce the metabolic normalisation. On the other hand, our results indicate that *Hippopodina* sp. presents a reduction in metabolic benefits of attaining a larger size at colder temperatures. Thus, for this encrusting species, one could predict that individuals living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction, for instance) while individuals living at lower temperatures will be under less pressure to allocate resources to size. Importantly, though, we note that there are many extrinsic factors that may affect abundance and selection on body size (e.g. predation, competition, etc; `r myCite('barneche2016procb')`). Yet, such differentials in the costs and benefits of attaining large size may help explain why some species living in cold temperatures present a slower pace of life (e.g. `r myCite('rosa2010jmba')`). Moreover, they may represent necessary improvements on current attempts to scale up energy from individuals to higher levels of organisation (e.g. `r myCite(c('allen2005','yvondurocher2012b','barneche2014ele','barneche2016procb'))`).

In the wild, partial predation is a common phenomenon experienced by colonial organisms, which must present different defence strategies `r myCite('sebens1987annurevecolsys')`. Our findings suggest that, for some species, predation might modulate the energetic demands of individuals of similar body mass in different environments (assuming that the size-manipulation approach herein adopted may serve as a surrogate for natural field predation; `r myCite('bone2005jembe')`), thus affecting life-history trajectories. For instance, at 25 ˚C, a 400-gram individual of *Bugula stolonifera* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Stolonifera'], 2)`-fold more energy than a similar-sized individual at 10 ˚C. Conversely, due to the estimated positive interaction between mass-scaling and temperature, a similar-sized individual of *B. neritina* will spend `r rounded(metabolicDiff$metabolicDiff[metabolicDiff$species == 'Bugula'], 2)`-fold more energy at 25 ˚C than at 10 ˚C. These differences, for instance, may help determine which species is a better competitor in different environments, as individual metabolic rates will also affect growth rates and time of reproduction onset `r myCite(c('sebens1987annurevecolsys', 'hart2009ecologyb'))`. It is worth noting that while we do not explicitly control for size manipulation in our study (i.e. evaluate the same model for non-manipulated individuals), previous evidence suggests that there is no effect of fragmentation on energetic demands of colonial organisms `r myCite('white2011amnat')`.

Here we have demonstrated that the effects of temperature interact in various ways with the mass-scaling of metabolic rates in different species of colonial marine invertebrates encompassing two distinct phyla. By adopting a size-manipulation approach, we achieved a more definitive test of mass-scaling of metabolic rates for cancelling the confounding effects of size and age, and by mimicking predation, which constitutes a fundamental ecological process in the wild. Mass-scaling exponents vary for different organisms, and are not consistent across growth forms nor phylogeny. These results suggest that more accurate predictions of abundance and biomass production at higher levels of organisation, from populations to ecosystems, will required a species-specific understanding of the interacting effects of mass scaling and temperature dependence, perhaps leading to improvements of current metabolic theories, or even the development of new ones.

**Acknowledgements**

Kearney, people who did the plates, ARC for much money (DP0987626, DP110101776, and whatever grants Dustin wants to add).

**Tables**

*Table 1.* Mass ranges and sampling efforts for the four species of Bryozoa and Porifera experimentally manipulated in the present study at two extreme temperatures.

*Table 2.* The best candidate model was chosen based on a selection procedure with nested models where non-significant fixed effects were successively removed based on leave-one-out cross-validation. Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models ($\widehat{elpd_{diff}}$).

**Figures**

*Figure 1.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*. Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively. *Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalisations (respectively), with bars over curves representing 95\% credible intervals, and numbers representing average estimates. *Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 10,000 MCMC iterations (translucent circles), with overlaid density heat-map. *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$. Regression fits were plotted using the average from the posterior distributions. In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior probability density curves (*n* = 512 values); differences between adjacent $x$ values (i.e. parameter estimates) were multiplied by the average of adjacent $y$ values (i.e. respective probability density values; $(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.

**References**